---
title: "COVID-19 Spatial Interventions Database"
author: "Priyanka Verma, Priscilla Ahn, Maxwell Enger, Noah Berkowitz"
date: "5/2/2020"
output: 
  ioslides_presentation: 
    highlight: null
    incremental: yes
    smaller: yes
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(covid19interventions)
library(shiny)
library(ggplot2)
library(tidyverse)
library(changepoint)
library(shinythemes)

data(mobility)
```

## Overview
- The United States confirmed its first case of COVID-19 on _Jan. 21, 2020_
- On _Feb. 26, 2020_, the first suspected case of a community spread in the U.S. was reported
- Local government bodies were first to enforce stay-at-home   to prevent the spread of COVID-19 before statewide measures were implemented

## Goal
- Identify counties that enforced stay-at-home measures before their state to address the COVID-19 pandemic 
- Understand if the county-level intervention to minimize direct human contact affects the curve

## Methodology
- We started by collecting data for the two most and least populated counties in every state- Slow process
- Google's Mobility dataset utilized to expedite the collection process

## Change in Mobility
- Movement information released by Google to support data-driven decision making in managing COVID-19
- Collected through Google products such as Google Maps
- Utilized to identify counties with the highest drop in workplace visits before statewide order implemented

## Change-Point Analysis
- Change-point detection technique used to determine change in workplace mobility curve before statewide stay-at-home order implemented for every county
- Specifically, At-Most-One-Change (AMOC) method used to determine a single point of where significant change is seen
- Mean of workplace mobility calculated pre and post change point
- Counties ranked based on mean change in workplace mobility post changepoint

## Change in Mobility by County
```{r, echo = FALSE}
fluidPage(
    sidebarLayout(
          sidebarPanel(
            htmlOutput("state_selector"),
            htmlOutput("county_selector"), width = "5"),
           mainPanel(plotOutput("cpplot"),
                     textOutput("pre_cp"),
                     textOutput("post_cp"),
                     width = 7)))
        

output$state_selector <- renderUI({
selectInput(
  inputId = "state", 
  label = "State:",
  choices = as.character(unique(df_joined$sub_region_1)),
  selected = "Alabama")
  })

output$county_selector <- renderUI({
  available <- df_joined[df_joined$sub_region_1 == input$state, "sub_region_2"]
  selectInput(
    inputId = "county", 
    label = "County:",
    choices = unique(available),
    selected = unique(available)[1])
})


  
  output$cpplot <- renderPlot({
    cp_calc <- df_joined %>% filter(sub_region_1 == input$state & 
                               sub_region_2 == input$county &
                               date < order_date &
                               !is.na(workplace_int)) %>%
    select(workplace_int) %>% 
    as.matrix() %>% 
    as.numeric() %>% 
    cpt.mean(method='AMOC')
    plot(cp_calc, cpt.col='blue', 
       ylab = "Workplace Mobility Change from Baseline (%)",
       xlab = "Days since February 15")
    output$pre_cp <- renderText({paste("Pre Changepoint Mean: ",
                                       round(param.est(cp_calc)$mean[1], 2))})
    output$post_cp <- renderText({paste("Post Changepoint Mean:",
                                        round(param.est(cp_calc)$mean[2]))})
    
})
```


## Interventions & COVID-19 Cases Curve
Overview of our database

## Before & During Interventions

## During & After Interventions

```{r}

```
## Rationale- NOAH 
-Covid-19 is a dynamic and easily spreadable virus. 
- Policy for COVID-19 has been decided on by governors in particular
- Expressing a Governor's political party will build overall trends in decision making for the US as a whole 
- The population is one of the largest factors in a factor in a pandemic.
      For example, New and NJ lead in terms of cases and are large population centers.(worldometers.info/coronavirus)
-Cases are the backbone for understanding the spread of disease. 
-Intervention types across the US were very different from Stay at home minors, to shelter in place
  - Interventions also occurred across both March of 2020 and April 2020
  
## Pop vs Intervention Date- NOAH




-This chart shows the intervention date with the total population of the state. N/A for the party is for territory without a governor.  
There is a clear difference in the political party's by the state for the state's intervention date. Democrats tend to be earlier by the republicans were later in March and early April.

-This chart was created using ggplot
## Political Party vs State Intervention Date- NOAH




-This chart shows the intervention type (policy type) with a political party. 
-Overall it shows the bipartisan policy between states for a stay at home order. It is also interesting to see a lack of shelter in place order between the states. 
-The chart shows the clear dominance geographically of the stay at home order.

-Note this table leaves off 9 states that have no SAH order (of any type)

-This chart was created using ggplot
## Flattening the Curve

## Future Work
```{r}

```



