---
title: "COVID-19 Interventions Analysis - Priscilla Ahn"
output: 
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{interventions_analysis_priscilla}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Introduction
Q. Do stay-at-home measures impact (flatten) the curve?
This R script seeks to locate the point in which COVID-19 cases begins to flatten
at the state- and county- level.

### Methodology
  1. "Linearize" exponential rise of cases using natural log\
  \
  <center> y = A$e^{kt}$ &rarr; ln(y) = kt + ln(A) </center>
  \
  2. Smooth natural log scale graph
  3. Calculate slope k between all the datepoints
  4. Locate breakpoints in k curve to find dates of greates slope change
  5. Calculate the "lag" between the SAH order date and the last breakpoint

### Data
This analysis requires cases data from the `covid19clark` package,
county-level data from our `covid19interventions` package, and state-level data
scraped from NYT, `sah`.

---
### Load data
```{r, warning=FALSE, message=FALSE, eval=FALSE}
# data load
library(covid19interventions)
# run if you don't have covid19clark
# devtools::install_github("agroimpacts/covid19clark", build_vignettes = TRUE)
library(covid19clark)
# covid interventions by county
data("county_interventions")
# covid cases
data("us_cases_daily")
# state SAH dates scraped from NY Times
data("sah")
```

### Sample data
```{r, warning=FALSE, message=FALSE, eval=FALSE}
# interventions
head(county_interventions)
# covid cases by county
head(us_cases_daily$county)
# nytimes sah dates
head(sah)
```

---

## Load packages
```{r, warning=FALSE, message=FALSE, eval=FALSE}
library(devtools)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lubridate)
library(strucchange)
library(segmented)
library(scales)
```

## Data prep
Cases need to be joined to state-level and county-level data
```{r, warning=FALSE, message=FALSE, eval=FALSE}
us_state <- us_cases_daily$state
us_county <- us_cases_daily$county

# state_join
tmptb <- sah %>% mutate(state = tolower(sah$state)) %>% 
  mutate(order_date = as.Date(order_date, format = "%m/%d/%Y")) %>% 
  filter(county == 'All') %>% select(state, order_date)
state_join <- us_state %>% select(state1, state2, date, cases, deaths) %>% 
  inner_join(tmptb, by = c("state1" = "state"))

# county_join
## drop words like "county" and "borough" from admin2
county_interventions$admin2 <- gsub("\\s*\\w*$", "", county_interventions$admin2)
county_interventions <- county_interventions %>% 
  mutate(acronym = tolower(county_interventions$acronym)) %>% 
  mutate(admin1 = tolower(county_interventions$admin1)) %>% 
  mutate(admin2 = tolower(county_interventions$admin2))
tmptb <- us_county %>% select(state1, county.x, date, cases, deaths)
county_join <- county_interventions %>%
  select(rank, acronym, admin1, admin2, SAH_County_Date, SAH_State_Date) %>% 
  inner_join(tmptb, by = c("admin1" = "state1", "admin2" = "county.x"))
## if SAH_County_Date is blank, fill with SAH_State_Date column
county_join$SAH_County_Date <- coalesce(county_join$SAH_County_Date,
                                        county_join$SAH_State_Date)

# function to raise case if needed
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}

# sample
head(state_join)
head(county_join)
```

# State-Level
Select the state of interest
```{r, warning=FALSE, message=FALSE, eval=FALSE}
# vector of state acronyms to loop
statename <- state_join %>% distinct(state1) %>% select(state1)

# placeholder 
statename <- statename %>% slice(1)
statename

# isolate variables of interest
state_tb <- state_join %>% select(state1, date, cases, order_date) %>% 
  filter(state1 == statename[[1]])

# convert date to julian day
state_tb$jday <- yday(state_tb$date)
```


## Calculate k
```{r, warning=FALSE, message=FALSE, eval=FALSE}
# calculate k value in y = Ae^(kx)
# normalize cases curve by exponential func using natural log
state_tb$cases_log <- log(state_tb$cases)
state_tb$k <- NA
state_tb$k[2:length(state_tb$k)] <- 
  diff(state_tb$cases_log)/diff(state_tb$jday)
state_tb <- state_tb %>% filter_all(all_vars(!is.infinite(.)))

# calculate fitted k value
## remove rows with any na values
state_tb_fit <- state_tb %>% drop_na()

## smooth cases_log
lw <- loess(cases_log ~ jday, state_tb_fit)
state_tb_fit$cases_logfit <- lw$fitted

## calculate k values to smoothed cases_log curve
state_tb_fit$kfit <- NA
state_tb_fit$kfit[2:length(state_tb_fit$kfit)] <- 
  diff(state_tb_fit$cases_logfit)/diff(state_tb_fit$jday)

# order_date
orderdate <- unique(state_tb$order_date)
orderjday <- yday(orderdate)
```

# Plot
```{r, fig.width=6, fig.height=4, fig.align = "center"}
p1 <- ggplot(state_tb, aes(x=date, y=cases)) +
  geom_point() +
  scale_x_date(date_breaks = "4 day", labels=date_format("%b-%d")) +
  labs(title = "COVID-19 Cases",
     subtitle = firstup(statename[[1]]),
     x = "Date", y = "Cases") +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.title = element_blank())
p1

p2 <- ggplot(state_tb_fit, aes(x=date, y=cases_log)) + geom_point() +
  geom_line(aes(x=date, y=cases_logfit), color = "#CD5C5C", size=1) +
  scale_x_date(date_breaks = "4 day", labels=date_format("%b-%d")) +
  labs(title = "COVID-19 Cases in Natural Log Scale",
     subtitle = firstup(statename[[1]]),
     x = "Date", y = "Cases (log scale)") +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.title = element_blank())
p2
```


# Locate breakpoints in cases data
```{r, warning=FALSE, message=FALSE, eval=FALSE}
# pull jday and kfit as numeric vectors
x <- state_tb_fit %>% pull(jday)
y <- state_tb_fit %>% pull(kfit)

# use intervention_breakpts function to find breakpoints within the data
output <- intervention_breakpts(x, y)
```

## Investigate outputs of `intervention_breakpts()`
```{r, warning=FALSE, message=FALSE, eval=FALSE}
## output 1: exported as a tibble
t <- output[[1]]

## output 2: linear fit to curve
my.lm <- output[[2]][[1]]
summary(my.lm)

## output 3: segments
my.seg <- output[[2]][[2]]
slope(my.seg)
summary(my.seg)

## output 4: modeled piecewise linear function
my.model <- output[[2]][[3]]

## output 5: estimated breakpoints in julian days
my.lines <- round(unname(output[[2]][[4]]))
```

## Plot outputs
```{r, fig.width=6, fig.height=4, fig.align = "center"}
# convert julian day to date
t_join <- t %>% left_join(my.model, by = c("jdates")) %>% 
  rename("k"="rate.x", "klinear"="rate.y")
t_join$date <- as.Date(t_join$jdates, origin = '2019-12-31')
my.lines_date <- as.Date(my.lines, origin = '2019-12-31')

# label vertical lines
vline1 <- data.frame(date = orderdate,
                    label = "SAH")
vline2 <- data.frame(date = my.lines_date,
                     label = "breakpoints")

# plot
p3 <- ggplot(t_join, aes(x=date, y=k)) + geom_point() +
  geom_line(aes(x=date, y=klinear)) +
  geom_vline(aes(xintercept=date, colour=label), data=vline1,
             linetype="dashed", size=1) +
  geom_vline(aes(xintercept=date, colour=label), data=vline2, size=1) +
  scale_x_date(date_breaks = "2 day", labels=date_format("%b-%d")) +
  scale_colour_manual(name = "Dates", values = c("breakpoints" = "grey",
                                               "SAH" = "#57B1C5")) +
  labs(title = "Breakpoints in k Curve",
     subtitle = firstup(statename[[1]]),
     caption = paste0("SAH Order Date = ", orderdate),
     x = "Date", y = "k") +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.title = element_blank())
p3
```

# Calculate lag
```{r, warning=FALSE, message=FALSE, eval=FALSE}
flatpoint_jday <- my.lines[length(my.lines)]
flatpoint_date <- as.Date(flatpoint_jday, origin = '2019-12-31')
lagdiff <- flatpoint_jday - orderjday
```

# Plot
```{r, fig.width=6, fig.height=4, fig.align = "center"}
# plot
df <- state_tb_fit %>% as.data.frame()
if (length(my.lines_date) == 1) {
  df_vline <- data.frame(dates = c(my.lines_date, flatpoint_date, orderdate),
                       labels = c("breakpoints", "flatten", "SAH"))
} else {
  df_vline <- data.frame(dates = c(my.lines_date[1:length(my.lines_date)-1],
                                   flatpoint_date, orderdate),
                         labels = c("breakpoints", "flatten", "SAH"))
  }

p4 <- ggplot(df, aes(x=date, y=cases_log)) + geom_point() +
  geom_line(aes(x=date, y=cases_logfit)) +
  geom_vline(aes(xintercept = dates, colour = labels), data = df_vline,
             linetype = 1, size = 1, show.legend = T) +
  scale_colour_manual(name = "Dates", values = c("breakpoints" = "grey",
                                                 "flatten" = "#E88147",
                                                 "SAH" = "#57B1C5")) +
  scale_x_date(date_breaks = "2 day", labels=date_format("%b-%d")) +
  labs(title = "Impact of SAH Order on Curve",
       subtitle = firstup(statename[[1]]),
       caption = paste0("Lag = ", lagdiff, " days"),
       x = "Date", y = "Cases (log scale)") +
  theme(axis.text.x = element_text(angle = 90))
p4
```

Together
```{r, fig.width=6, fig.height=4, fig.align = "center"}
grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2)
```

- County level
```{r, warning=FALSE, message=FALSE, eval=FALSE}
# vector of state acronyms to loop
countyname <- county_join %>% distinct(admin1, admin2)

# temp before I figure out looping
countyname <- countyname %>% slice(1)
countyname

# isolate variables of interest
county_tb <- county_join %>%
  select(admin1, admin2, date, cases, SAH_County_Date) %>% 
  filter(admin1 == countyname[[1]] & admin2 == countyname[[2]])

# convert date to julian day
county_tb$jday <- yday(county_tb$date)

# calculate k value in y = Ae^(kx)
# normalize cases curve by exponential func using natural log
county_tb$cases_log <- log(county_tb$cases)
county_tb$k <- NA
county_tb$k[2:length(county_tb$k)] <- 
  diff(county_tb$cases_log)/diff(county_tb$jday)
county_tb <- county_tb %>% filter_all(all_vars(!is.infinite(.)))

# calculate fitted k value
## remove rows with any na values
county_tb_fit <- county_tb %>% drop_na()

## smooth cases_log
lw <- loess(cases_log ~ jday, county_tb_fit)
county_tb_fit$cases_logfit <- lw$fitted

## calculate k values to smoothed cases_log curve
county_tb_fit$kfit <- NA
county_tb_fit$kfit[2:length(county_tb_fit$kfit)] <- 
  diff(county_tb_fit$cases_logfit)/diff(county_tb_fit$jday)

# order_date
orderdate <- unique(county_tb$SAH_County_Date)
orderjday <- yday(orderdate)

# pull jday and kfit as numeric vectors
x <- county_tb_fit %>% pull(jday)
y <- county_tb_fit %>% pull(kfit)

# use intervention_breakpts function to find breakpoints within the data
output <- intervention_breakpts(x, y)

## outputs
t <- output[[1]]
my.lm <- output[[2]][[1]]
summary(my.lm)
my.seg <- output[[2]][[2]]
slope(my.seg)
summary(my.seg)
my.model <- output[[2]][[3]]
my.lines <- round(unname(output[[2]][[4]]))

# calculate lag
flatpoint_jday <- my.lines[length(my.lines)]
flatpoint_date <- as.Date(flatpoint_jday, origin = '2019-12-31')
lagdiff <- flatpoint_jday - orderjday

# plot 1
p1 <- ggplot(county_tb, aes(x=date, y=cases)) +
  geom_point() +
  scale_x_date(date_breaks = "4 day", labels=date_format("%b-%d")) +
  labs(title = "COVID-19 Cases",
     subtitle = paste0(firstup(countyname[[2]]), ", ",
                       firstup(countyname[[1]])),
     x = "Date", y = "Cases") +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.title = element_blank())
p1

# plot 2
p2 <- ggplot(county_tb_fit, aes(x=date, y=cases_log)) +
  geom_point() +
  geom_line(aes(x=date, y=cases_logfit), color = "#CD5C5C", size=1) +
  scale_x_date(date_breaks = "4 day", labels=date_format("%b-%d")) +
  labs(title = "COVID-19 Cases in Natural Log Scale",
     subtitle = paste0(firstup(countyname[[2]]), ", ",
                       firstup(countyname[[1]])),
     x = "Date", y = "Cases (log scale)") +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.title = element_blank())
p2

# plot 3
## convert julian day to date
t_join <- t %>% left_join(my.model, by = c("jdates")) %>% 
  rename("k"="rate.x", "klinear"="rate.y")
t_join$date <- as.Date(t_join$jdates, origin = '2019-12-31')
my.lines_date <- as.Date(my.lines, origin = '2019-12-31')

## label vertical lines
vline1 <- data.frame(date = orderdate,
                    label = "SAH")
vline2 <- data.frame(date = my.lines_date,
                     label = "breakpoints")

## plot
p3 <- ggplot(t_join, aes(x=date, y=k)) + geom_point() +
  geom_line(aes(x=date, y=klinear)) +
  geom_vline(aes(xintercept=date, colour=label), data=vline1,
             linetype="dashed", size=1) +
  geom_vline(aes(xintercept=date, colour=label), data=vline2, size=1) +
  scale_x_date(date_breaks = "2 day", labels=date_format("%b-%d")) +
  scale_colour_manual(name = "Dates", values = c("breakpoints" = "grey",
                                               "SAH" = "#57B1C5")) +
  labs(title = "Breakpoints in k Curve",
     subtitle = paste0(firstup(countyname[[2]]), ", ",
                       firstup(countyname[[1]])),
     caption = paste0("SAH Order Date = ", orderdate),
     x = "Date", y = "k") +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.title = element_blank())
p3

# plot 4
df <- county_tb_fit %>% as.data.frame()
if (length(my.lines_date) == 1) {
  df_vline <- data.frame(dates = c(my.lines_date, flatpoint_date, orderdate),
                       labels = c("breakpoints", "flatten", "SAH"))
} else {
  df_vline <- data.frame(dates = c(my.lines_date[1:length(my.lines_date)-1],
                                   flatpoint_date, orderdate),
                         labels = c("breakpoints", "flatten", "SAH"))
  }

p4 <- ggplot(county_tb_fit, aes(x=date, y=cases_log)) + geom_point() +
  geom_line(aes(x=date, y=cases_logfit)) +
  geom_vline(aes(xintercept = dates, colour = labels), data = df_vline,
             linetype = 1, size = 1, show.legend = T) +
  scale_colour_manual(name = "Dates", values = c("breakpoints" = "grey",
                                                 "flatten" = "#E88147",
                                                 "SAH" = "#57B1C5")) +
  scale_x_date(date_breaks = "2 day", labels=date_format("%b-%d")) +
  labs(title = "Did SAH order flatten the curve?",
       subtitle = paste0(firstup(countyname[[2]]), ", ",
                         firstup(countyname[[1]])),
       caption = paste0("Lag = ", lagdiff, " days"),
       x = "Date", y = "Cases (log scale)") +
  theme(axis.text.x = element_text(angle = 90))
p4

grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2)
```

# elbow point
```{r, warning=FALSE, message=FALSE}
# elbowPoint(jdates, k)
```






