---
title: "COVID-19 Spatial Interventions Project"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{COVID-19 Spatial Interventions Project Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(covid19interventions)
```

## Goal
Compile and analyze COVID-19 related spatial interventions in the U.S. at the county level

## Data
The data will be extracted from various sources and added to the COVID-19 US Spatial Response database. We are currently looking into putting together a list of verified twitter accounts for US counties. 

## Responsibilities
**Priyanka**: Research & data collection<br/>
**Noah**: Research & data collection <br/>
**Priscilla**: Research & analysis<br/>
**Max**: Research & analysis<br/>

## Timeline
- **Week 1**: Identify data sources
- **Week 2**: Write scripts to extract data
- **Week 3**: Analyze data
